\textbf{COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata}

### 2023-03-23 manuscript_v2/01-background.qmd
code for age analysis

The vaccines against COVID-19 have been considered a major achievement of modern medicine as their rapid development[@davis2022] allowed the start of broad vaccination campaigns towards the end of 2020 in certain countries such as the US[@tanne2020] and Canada[@bogoch2022]. This made some believe that vaccines were destined to be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, although it has been estimated that COVID-19 vaccines have prevented around 14 million of deaths worldwide [@watson2022], their implementation has been far from being equal to that of the vaccines of smallpox and polio, which were implemented on a global scale and that were indeed crucial to control these diseases [@kayser2021]. In fact, the rollout of COVID-19 vaccines has faced multiple challenges since its inception which ultimately have hampered the goal of global immunity.

This ongoing lack of a successful global rollout of the COVID-19 vaccines is a multifaceted issue resulting from, as among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020]. In particular, it is well established that differences in vaccination uptake have been present even in the case of countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for lower vaccine uptake in these cases have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020]. More specifically, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of a visible minorities or Indigenous, and those that are financially unstable [@guay2022;@muhajarine2021;@hussain2022].

However, it is important to consider that vaccination uptake can also be influenced by geographical (spatial) factors. 
In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021]. Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022]. Overall, the existing body of  research on geographical/spatial differences in vaccination has provided valuable information that can be used by decision-makers in the development of public health policies that are focused on improving vaccination within specific administrative or geographical areas. However, spatial analyses of vaccination have been very limited in Canada, as existing studies have been focused only at certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022]. However, there is a lack of studies that explore spatial heterogeneities in vaccination at the intra-provincial level.

Ontario is the most populated province in Canada, and has recently seen major changes on its healthcare structure as between 2006 and 2019, it provided healthcare access using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs). However, this approach was complex and caused systemic inequalities [@tsasis2012]. In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province[@dong2022]. Because the adoption of the Health Regions has been recent, there is an ongoing need to analyze the impact of their adoption and intra-regional differences that might exist, which is specially important in the context of the COVID-19 pandemic.

To address this need, in this study we analyzed differences in self-reported COVID-19 vaccination status using socio-economic (e.g., income, racial/ethnic identification), and spatial  information at the level of the Health Regions, to identify the existence of intra-provincial differences that might need to be addressed to ensure that the healthcare system of Ontario is more inclusive and that responds to the needs of its most vulnerable population.

>>>>

The vaccines against COVID-19 have been considered a major achievement of modern medicine as their rapid development allowed the start of broad vaccination campaigns towards the end of 2020 in certain countries, such as the US and Canada [@davis2022;@bogoch2022;@tanne2020]. This made some believe that vaccines were destined to be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, although it has been estimated that COVID-19 vaccines have prevented around 14 million of deaths worldwide [@watson2022], their implementation has been far from being equal to that of the vaccines of smallpox and polio, which were implemented on a global scale and that were indeed crucial to control these diseases [@kayser2021]. In fact, the rollout of COVID-19 vaccines has faced multiple challenges since its inception which ultimately have hampered their use to achieve the ultimate goal of global immunity.

This problematic of COVID-19 vaccines rollout is a multifaceted issue resulting from, among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020]. In particular, it is well established that differences in vaccination uptake have been present even in countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for lower vaccine uptake in these cases have included medical mistrust due to systemic medical racism [@stoler2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020]. Moreover, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of a visible minorities or Indigenous, and those that are financially unstable [@guay2022;@muhajarine2021;@hussain2022].

However, it is important to consider that vaccination uptake can also be influenced by geographical (spatial) factors. 
In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021]. Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022]. Overall, the evidence provided by the literature demonstrates the existence of spatially-driven heterogeneities in vaccine uptake that be used by decision-makers in the development of public health policies that are focused on addressing these disparities within specific administrative or geographical areas.

However, such analyses have been carried mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022]. Thus, there is a need for studies that explore spatial differences in vaccination within the Canadian territory and that consequently, can help identify disparities that need to be addressed within specific areas in each province.

This need is particularly important in the case of Ontario, the most populated province in Canada. Between 2006 and 2019, Ontario provided healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs). However, this approach was complex, bureaucratic, and led to systemic inequalities [@tsasis2012]. In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province [@dong2022]. Because the adoption of the Health Regions occurred at a relatively recent time, there is an ongoing need to analyze the impact of this measure and identify the existence of intra-regional differences that might exist, and which could be specially important in the context of the COVID-19 pandemic.

In this study, we analyzed differences in self-reported COVID-19 vaccination status in Ontario using socio-economic (e.g., income, racial/ethnic identification), and spatial  information at the level of the Health Regions to identify the existence of differences that might need to be addressed to ensure that the healthcare system of Ontario is more inclusive and that responds to the needs of its most vulnerable population.

### 2023-03-23 manuscript_v2/02-methods.qmd
code for age analysis

Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions. The low representation of these Health Regions in the dataset is noticeable in @fig-map, where the number of cities in each Health Region is relatively low compared to the most populated Health Regions, such as the Toronto or Central Regions. We omitted the North East and North East Health Regions from further analyses due to the low number of entries. Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions.

@@

<!-- $$ -->
<!-- \begin{aligned} -->
<!-- \log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Race)} + \beta_6 \textrm{ (Income} \times \textrm{Race)} -->
<!-- \end{aligned} -->
<!-- $$ {#eq-model3} -->

<!-- Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a COVID-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term.  -->

>>>>

Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions. The low representation of these Health Regions in the dataset is noticeable in @fig-map,which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions. We omitted the North East and North East Health Regions from further analyses due to the low number of entries. Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions.

@@

### 2023-03-23 manuscript_v2/03-results.qmd
code for age analysis

@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries from the survey in the dataset after cleaning was 3,551. Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it. Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\% of the total number of entries, those within the CAD 25,000-59,999 income bracket represented 25% of the total sample, and those with an income above CAD 60,000 represented 38 \% of the sample; across all income brackets, the percentage of individuals that reported having received a first dose of the vaccine was consistent, above 69\%.

>>>>

@tbl-descriptive-stats shows the descriptive statistics from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries analyzed was 3,551. Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it. Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\% of the total number of entries, those within the CAD 25,000-59,999 income bracket represented 25% of the total sample, and those with an income above CAD 60,000 represented 38 \% of the sample; across all income brackets, the percentage of individuals that reported having received a first dose of the vaccine was consistent, above 69\%.

### 2023-03-23 manuscript_v2/main.tex
code for age analysis

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

The vaccines against COVID-19 have been considered a major achievement
of modern medicine as their rapid
development\textsuperscript{\protect\hyperlink{ref-davis2022}{1}}
allowed the start of broad vaccination campaigns towards the end of 2020
in certain countries such as the
US\textsuperscript{\protect\hyperlink{ref-tanne2020}{2}} and
Canada\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3}}. This made
some believe that vaccines were destined to be a determinant factor in a
rapid ending of the
pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
However, although it has been estimated that COVID-19 vaccines have
prevented around 14 million of deaths
worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
implementation has been far from being equal to that of the vaccines of
smallpox and polio, which were implemented on a global scale and that
were indeed crucial to control these
diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
fact, the rollout of COVID-19 vaccines has faced multiple challenges
since its inception which ultimately have hampered the goal of global
immunity.

This ongoing lack of a successful global rollout of the COVID-19
vaccines is a multifaceted issue resulting from, as among other things,
the development of new variants due to inadequate public health
measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
in vaccine access between high- and low-income
countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{8},\protect\hyperlink{ref-tamey2022}{9}},
vaccine
hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{10}}, and
differences in vaccination uptake across different segments of the
population\textsuperscript{\protect\hyperlink{ref-malik2020}{11}}. In
particular, it is well established that differences in vaccination
uptake have been present even in the case of countries that have had
ample access to vaccines since 2020 (such as the US, the UK, and
Canada), where lower vaccine uptake has been observed within racial
minorities (i.e., individuals that identify as Black, Asian, or
Indigenous), and in individuals within low income
brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
Reasons given for lower vaccine uptake in these cases have included
medical mistrust due to systemic medical racism, mistrust in vaccines,
and the influence of conspiracy
theories\textsuperscript{\protect\hyperlink{ref-willis2021}{12},\protect\hyperlink{ref-stoler2021}{14},\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
More specifically, in the case of Canada, lower vaccine uptake has been
observed in young individuals, those with a low educational level,
households with children, those without a regular healthcare provider,
individuals that identify as part of a visible minorities or Indigenous,
and those that are financially
unstable\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.

However, it is important to consider that vaccination uptake can also be
influenced by geographical (spatial) factors. In this regard,
differences in COVID-19 vaccination rates have been associated with
varied regional attitudes towards
vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{11}},
spatial differences in vaccine access and supply, vaccination location
availability, and lack of prioritization of areas where vulnerable
groups
reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3},\protect\hyperlink{ref-nguyen2021}{22}}.
Other studies have also shown heterogeneity in vaccine uptake within
small governmental administrative units such as
counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{23}--\protect\hyperlink{ref-bhuiyan2022}{26}},
and that and that accounting for geographical differences in vaccination
can help predict patterns of booster
uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
the existing body of research on geographical/spatial differences in
vaccination has provided valuable information that can be used by
decision-makers in the development of public health policies that are
focused on improving vaccination within specific administrative or
geographical areas. However, spatial analyses of vaccination have been
very limited in Canada, as existing studies have been focused only at
certain cities (such as
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}}, or
Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}), or
have explored differences at a province-wide
level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. However,
there is a lack of studies that explore spatial heterogeneities in
vaccination at the intra-provincial level.

Ontario is the most populated province in Canada, and has recently seen
major changes on its healthcare structure as between 2006 and 2019, it
provided healthcare access using 14 intra-provincial divisions called
the Local Health Integrated Networks (LHINs). However, this approach was
complex and caused systemic
inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
late 2019, the 14 LHINs were phased out and the areas they covered were
incorporated into 6 Health Regions (North East, North West, Central,
Toronto, West, and East) in an effort to improve the healthcare system
of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
Because the adoption of the Health Regions has been recent, there is an
ongoing need to analyze the impact of their adoption and intra-regional
differences that might exist, which is specially important in the
context of the COVID-19 pandemic.

To address this need, in this study we analyzed differences in
self-reported COVID-19 vaccination status using socio-economic (e.g.,
income, racial/ethnic identification), and spatial information at the
level of the Health Regions, to identify the existence of
intra-provincial differences that might need to be addressed to ensure
that the healthcare system of Ontario is more inclusive and that
responds to the needs of its most vulnerable population.

@@

Following an assessment of the number of entries corresponding to each
Health Region in the final dataset, only 107 observations (4.3\% of the
total) corresponded to cities located in the North West and North East
Health Regions. The low representation of these Health Regions in the
dataset is noticeable in Figure~\ref{fig-map}, where the number of
cities in each Health Region is relatively low compared to the most
populated Health Regions, such as the Toronto or Central Regions. We
omitted the North East and North East Health Regions from further
analyses due to the low number of entries. Therefore, the total number
of unique entries used for analysis was 3,551 which included the East,
Central, Toronto, and West Health Regions.

@@

Table~\ref{tbl-descriptive-stats} shows the descriptive statistics
(uncorrected) from the Fields COVID-19 survey data for vaccination
status and each of the covariates analyzed. The total number of entries
from the survey in the dataset after cleaning was 3,551. Overall, 26.9\%
of survey respondents (958) reported not having received the first dose
of the vaccine, whereas 73.1\% (2,593) reported having received it.
Within each socio-economic factor, respondents who reported living in a
household with an income under CAD 25,000 represented 37\% of the total
number of entries, those within the CAD 25,000-59,999 income bracket
represented 25\% of the total sample, and those with an income above CAD
60,000 represented 38 \% of the sample; across all income brackets, the
percentage of individuals that reported having received a first dose of
the vaccine was consistent, above 69\%.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
2020:m4799.
doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}

\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
social and policy view of vaccination efforts in ontario, canada.
\emph{Cambridge Journal of Regions, Economy and Society}. Published
online November 2022.
doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}

>>>>

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

The vaccines against COVID-19 have been considered a major achievement
of modern medicine as their rapid development allowed the start of broad
vaccination campaigns towards the end of 2020 in certain countries, such
as the US and
Canada\textsuperscript{\protect\hyperlink{ref-davis2022}{1}--\protect\hyperlink{ref-tanne2020}{3}}.
This made some believe that vaccines were destined to be a determinant
factor in a rapid ending of the
pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
However, although it has been estimated that COVID-19 vaccines have
prevented around 14 million of deaths
worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
implementation has been far from being equal to that of the vaccines of
smallpox and polio, which were implemented on a global scale and that
were indeed crucial to control these
diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
fact, the rollout of COVID-19 vaccines has faced multiple challenges
since its inception which ultimately have hampered their use to achieve
the ultimate goal of global immunity.

This problematic of COVID-19 vaccines rollout is a multifaceted issue
resulting from, among other things, the development of new variants due
to inadequate public health
measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
in vaccine access between high- and low-income
countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{8},\protect\hyperlink{ref-tamey2022}{9}},
vaccine
hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{10}}, and
differences in vaccination uptake across different segments of the
population\textsuperscript{\protect\hyperlink{ref-malik2020}{11}}. In
particular, it is well established that differences in vaccination
uptake have been present even in countries that have had ample access to
vaccines since 2020 (such as the US, the UK, and Canada), where lower
vaccine uptake has been observed within racial minorities (i.e.,
individuals that identify as Black, Asian, or Indigenous), and in
individuals within low income
brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
Reasons given for lower vaccine uptake in these cases have included
medical mistrust due to systemic medical
racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{14}}, mistrust
in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{12}}, and
the influence of conspiracy
theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
Moreover, in the case of Canada, lower vaccine uptake has been observed
in young individuals, those with a low educational level, households
with children, those without a regular healthcare provider, individuals
that identify as part of a visible minorities or Indigenous, and those
that are financially
unstable\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.

However, it is important to consider that vaccination uptake can also be
influenced by geographical (spatial) factors. In this regard,
differences in COVID-19 vaccination rates have been associated with
varied regional attitudes towards
vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{11}},
spatial differences in vaccine access and supply, vaccination location
availability, and lack of prioritization of areas where vulnerable
groups
reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2},\protect\hyperlink{ref-nguyen2021}{22}}.
Other studies have also shown heterogeneity in vaccine uptake within
small governmental administrative units such as
counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{23}--\protect\hyperlink{ref-bhuiyan2022}{26}},
and that and that accounting for geographical differences in vaccination
can help predict patterns of booster
uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
the evidence provided by the literature demonstrates the existence of
spatially-driven heterogeneities in vaccine uptake that be used by
decision-makers in the development of public health policies that are
focused on addressing these disparities within specific administrative
or geographical areas.

However, such analyses have been carried mostly in territories outside
of Canada, where available studies have been focused in certain cities
(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}},
or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}),
or have explored differences at a province-wide
level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Thus, there
is a need for studies that explore spatial differences in vaccination
within the Canadian territory and that consequently, can help identify
disparities that need to be addressed within specific areas in each
province.

This need is particularly important in the case of Ontario, the most
populated province in Canada. Between 2006 and 2019, Ontario provided
healthcare access to its inhabitants using 14 intra-provincial divisions
called the Local Health Integrated Networks (LHINs). However, this
approach was complex, bureaucratic, and led to systemic
inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
late 2019, the 14 LHINs were phased out and the areas they covered were
incorporated into 6 Health Regions (North East, North West, Central,
Toronto, West, and East) in an effort to improve the healthcare system
of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
Because the adoption of the Health Regions occurred at a relatively
recent time, there is an ongoing need to analyze the impact of this
measure and identify the existence of intra-regional differences that
might exist, and which could be specially important in the context of
the COVID-19 pandemic.

In this study, we analyzed differences in self-reported COVID-19
vaccination status in Ontario using socio-economic (e.g., income,
racial/ethnic identification), and spatial information at the level of
the Health Regions to identify the existence of differences that might
need to be addressed to ensure that the healthcare system of Ontario is
more inclusive and that responds to the needs of its most vulnerable
population.

@@

Following an assessment of the number of entries corresponding to each
Health Region in the final dataset, only 107 observations (4.3\% of the
total) corresponded to cities located in the North West and North East
Health Regions. The low representation of these Health Regions in the
dataset is noticeable in Figure~\ref{fig-map},which shows that responses
from these areas came from a relatively low number of cities when
compared to the most populated Health Regions, such as the Toronto or
Central Regions. We omitted the North East and North East Health Regions
from further analyses due to the low number of entries. Therefore, the
total number of unique entries used for analysis was 3,551 which
included the East, Central, Toronto, and West Health Regions.

@@

Table~\ref{tbl-descriptive-stats} shows the descriptive statistics from
the Fields COVID-19 survey data for vaccination status and each of the
covariates analyzed. The total number of entries analyzed was 3,551.
Overall, 26.9\% of survey respondents (958) reported not having received
the first dose of the vaccine, whereas 73.1\% (2,593) reported having
received it. Within each socio-economic factor, respondents who reported
living in a household with an income under CAD 25,000 represented 37\%
of the total number of entries, those within the CAD 25,000-59,999
income bracket represented 25\% of the total sample, and those with an
income above CAD 60,000 represented 38 \% of the sample; across all
income brackets, the percentage of individuals that reported having
received a first dose of the vaccine was consistent, above 69\%.

@@

\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
social and policy view of vaccination efforts in ontario, canada.
\emph{Cambridge Journal of Regions, Economy and Society}. Published
online November 2022.
doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}

\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
2020:m4799.
doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}
